---
title: "Single- and multi-variate survival analysis"
author: "Zhiyuan Hu"
date: "15/05/2020 (last modified: `r Sys.Date()`)"
output: 
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---

```{r setup, include=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = FALSE, fig.width = 6, fig.height = 4, warning = F)

# devtools::install_github('shenorrlab/bseq-sc')
# devtools::install_github('shenorrlab/csSAM')
library(bseqsc)

library(MASS)
library(xbioc)
library(e1071) #svm
library(preprocessCore)

library(survminer)
library(survival)
# library(powerSurvEpi)
library(Biobase)
# BiocManager::install("SingleCellExperiment")
# BiocManager::install("scater")
# library(SingleCellExperiment)
# library(scater)

library(ggbeeswarm)
library(ggrepel)
library(viridis)
library(ggplot2)
library(cowplot)
library(ggpubr)

# library(edgeR)
# library(limma)
# library(copynumber)

source("codings/modified_functions20190204.R")
```

# Summary 

This script aims to perform deconvolution & survival analysis on the HH-cohort and visualise the expression data as well as the deconvolution results.

# Load data

```{r}
# sig_matrix <- readRDS("../../Project scRNAseq analysis/A_SmartSeq2/man_analysis2_20180629/R_deconvolution/rds/20190213sig_matrix_new.rds")

sig_matrix_new <-  read.csv("../TableS7_selected_signature_matrix_52genes20190204.csv", as.is = T, row.names = 1)
sig_matrix_new <- as.matrix(sig_matrix_new)
```


```{r}
counts <- read.csv("../HH001-150_NormalizedData_cleaned.csv", as.is = T, row.names = 1)

rownames(counts)[7] <- "MLF1IP"

colnames(counts) <- gsub(pattern = "HH", replacement = "HH-", x = colnames(counts))
colnames(counts) <- gsub(pattern = "[.]", replacement = "", x = colnames(counts))
counts[1:5,1:5]
# counts <- read.csv("../HH_cohort/HH001-150_Rawdata_cleaned.csv", as.is = T, row.names = 1)
# counts[1:5,1:5]
```



```{r}
clic_dt <- readxl::read_xlsx("../clinical_data.xlsx")
colnames(clic_dt) <- gsub(pattern = " ", replacement = "_", x = colnames(clic_dt))
clic_dt$Overall_survival <- as.numeric(clic_dt$Overall_survival)
clic_dt$Event <- as.numeric(clic_dt$Event)

clic_dt$Residual_disease <- clic_dt$`Residual_disease_0_optimal_1_sub-optimal`
clic_dt$Residual_disease <- as.factor(as.numeric(clic_dt$Residual_disease))

head(clic_dt)
```

```{r}
sum(rownames(counts) %in% rownames(sig_matrix_new))
```

# Deconvolution

```{r, fig.height=10}
cibersort_fit <- cibersort_my(sig_matrix = sig_matrix_new, exprs_unlogged = counts)
# cibersort_fit2 <- cibersort_my(sig_matrix = sig_matrix, exprs_unlogged = counts)

# write.csv(nanostring_fit, file = "results/cibersort_HHcohort150_output.csv")
colnames(cibersort_fit)[c(1,2,4,5)] <- c("Differentiated","KRT17","Cell_cycle","Ciliated")

clic_dt <- cbind(clic_dt, cibersort_fit[match(clic_dt$Anonymised_code, rownames(cibersort_fit)),])
# write.csv(clic_dt, "data/cdlinical_data_unfiltered.csv")
```

# Filtering

```{r}
table(clic_dt$EMT < 0.95)
# FALSE  TRUE 
#     2   148 
clic_filtered <- clic_dt[clic_dt$EMT < 0.95,]
dim(clic_filtered)

table(clic_filtered$Summary_histology) 
clic_filtered <- clic_filtered[clic_filtered$Summary_histology != "Endometrial serous carcinoma",] # remove Endometrial serous carcinoma
dim(clic_filtered)

clic_filtered$Event[clic_filtered$Overall_survival == 29] <- 0 # die from compliments; 1 patient
dim(clic_filtered)

sum(is.na(clic_filtered$Overall_survival)) # 6 dont have survival info
clic_filtered <- clic_filtered[!is.na(clic_filtered$Overall_survival),]
dim(clic_filtered)
```


## Single-variate analysis

```{r}
surv_fit <- coxph(Surv(Overall_survival, Event)~EMT, data=as.data.frame(clic_dt))
summary(surv_fit)
```

```{r}
surv_fit <- coxph(Surv(Overall_survival, Event)~EMT, data=as.data.frame(clic_filtered))
summary(surv_fit)
```



```{r}
surv_fit <- coxph(Surv(Overall_survival, Event)~Summary_stage, data=as.data.frame(clic_filtered))
summary(surv_fit)
```

```{r}
surv_fit <- coxph(Surv(Overall_survival, Event)~Summary_grade, data=as.data.frame(clic_filtered))
summary(surv_fit)
```

```{r}
surv_fit <- coxph(Surv(Overall_survival, Event)~Residual_disease, data=as.data.frame(clic_filtered))
summary(surv_fit)
```

## Dual-variate analysis

```{r}
surv_fit <- coxph(Surv(Overall_survival, Event)~EMT+Summary_stage, data=as.data.frame(clic_filtered))
summary(surv_fit)
```

```{r}
surv_fit <- coxph(Surv(Overall_survival, Event)~EMT+Summary_grade, data=as.data.frame(clic_filtered))
summary(surv_fit)
```

```{r}
surv_fit <- coxph(Surv(Overall_survival, Event)~EMT+Residual_disease, data=as.data.frame(clic_filtered))
summary(surv_fit)
```

## Multiple-variate analysis

```{r multivariate}
surv_fit <- coxph(Surv(Overall_survival, Event)~EMT+Summary_grade+Summary_stage, data=as.data.frame(clic_filtered))
summary(surv_fit)
```

```{r}
# xlsx::write.xlsx(clic_filtered,"results/clincData_deconRes_filtered20191106.xlsx")
```


```{r multivariate-res-table}
res_tb <- data.frame(Variate = c("EMT","Grade","Stage"),
                     Hazard_ratio = summary(surv_fit)$coefficients[,2],
                     p = summary(surv_fit)$coefficients[,5],
                     CI_low = summary(surv_fit)$conf.int[,3],
                     CI_high = summary(surv_fit)$conf.int[,4])
res_tb$Hazard_ratio[2] <- 1/res_tb$Hazard_ratio[2]
res_tb[2, 4:5] <- 1/res_tb[2,c(5,4)]
# write.csv(res_tb, "results/survival_Fitres_HH20191106.csv", row.names = F)
knitr::kable(res_tb, row.names = F)
```


```{r multivariate-new}
clic_filtered$`Residual_disease_0_optimal_1_sub-optimal` <- as.numeric(clic_filtered$`Residual_disease_0_optimal_1_sub-optimal`)
surv_fit <- coxph(Surv(Overall_survival, Event)~EMT+Summary_grade+Summary_stage+ `Residual_disease_0_optimal_1_sub-optimal`, data=as.data.frame(clic_filtered)[,])
summary(surv_fit)
```

```{r multivariate-res-table2}
res_tb <- data.frame(Variate = c("EMT","Grade","Stage","Residual_disease"),
                     Hazard_ratio = summary(surv_fit)$coefficients[,2],
                     p = summary(surv_fit)$coefficients[,5],
                     CI_low = summary(surv_fit)$conf.int[,3],
                     CI_high = summary(surv_fit)$conf.int[,4])
res_tb$Hazard_ratio[2] <- 1/res_tb$Hazard_ratio[2]
res_tb[2, 4:5] <- 1/res_tb[2,c(5,4)]
# write.csv(res_tb, "results/survival_Fitres_with_residualDisease_HH20200515.csv", row.names = F)
knitr::kable(res_tb)
```

# Association between EMT and residual disease

suboptimal group vs optimal group

```{r}
t.test(clic_filtered$EMT[clic_filtered$Residual_disease == 0], ## suboptimal
       clic_filtered$EMT[clic_filtered$Residual_disease == 1]) ## optimal
```

suboptimal late-stage group vs optimal late-stage group

```{r}
t.test(clic_filtered$EMT[clic_filtered$Residual_disease == 0 & clic_filtered$Summary_stage == "late"], ## suboptimal
       clic_filtered$EMT[clic_filtered$Residual_disease == 1 & clic_filtered$Summary_stage == "late"]) ## optimal
```

# Visualisation

## Stacked barplot

```{r Stacked barplot-preparedata}
svr_fit_oxo <- clic_dt[,26:30]
svr_fit_new_diff <- svr_fit_oxo[svr_fit_oxo[,1] > 0.5,]
svr_fit_new_diff <- svr_fit_new_diff[order(svr_fit_new_diff[,1], decreasing = F),]
svr_fit_new_krt <- svr_fit_oxo[svr_fit_oxo[,2] > 0.5,]
svr_fit_new_krt <- svr_fit_new_krt[order(svr_fit_new_krt[,2], decreasing = F),]
svr_fit_new_emt <- svr_fit_oxo[svr_fit_oxo[,3] > 0.5,]
svr_fit_new_emt <- svr_fit_new_emt[order(svr_fit_new_emt[,3], decreasing = F),]
svr_fit_new_cc <- svr_fit_oxo[svr_fit_oxo[,4] > 0.5,]
svr_fit_new_cc <- svr_fit_new_cc[order(svr_fit_new_cc[,4], decreasing = F),]

svr_fit_new_others <- svr_fit_oxo[!rownames(svr_fit_oxo) %in% c(rownames(svr_fit_new_diff),rownames(svr_fit_new_krt),
                                                               rownames(svr_fit_new_emt),rownames(svr_fit_new_cc)),]

plot.data <- data.frame(sample = rep(rownames(svr_fit_oxo), 5),
                        state = rep(c("Differentiated","KRT17","EMT","Cell cycle","Ciliated"), 
                                    each = nrow(svr_fit_oxo)),
                        score = c(svr_fit_oxo[,1],
                                  svr_fit_oxo[,2],
                                  svr_fit_oxo[,3],
                                  svr_fit_oxo[,4],
                                  svr_fit_oxo[,5]))
plot.data$sample <- factor(plot.data$sample, levels = c(rownames(svr_fit_new_diff),rownames(svr_fit_new_krt),
                                                               rownames(svr_fit_new_emt),rownames(svr_fit_new_cc),
                                                        rownames(svr_fit_new_others)))
plot.data$state <- factor(plot.data$state, levels = c("Differentiated","KRT17","EMT","Cell cycle","Ciliated"))
# differentiated #A16BA3
# F6A000 KRT7
# FFDE00 EMT
# CC #4DD9FF
#F3A1B6 Ciliated
```

```{r, fig.width=10, fig.height=3}
ggplot(plot.data) + geom_bar(aes(y = score, x = sample, fill = state),
                           stat="identity") + theme_classic() + 
  scale_fill_manual(values = c("#A16BA3","#F6A000","#FFDE00","#4DD9FF","#F3A1B6"), 
                    breaks = c("Differentiated","KRT17","EMT","Cell cycle","Ciliated"), 
                    labels = c("Differentiated","KRT17","EMT","Cell cycle","Ciliated")) +
  xlab("Bulk tumour samples \n (each column is one tumour case)") + ylab("Proportions of cell states") +
  theme(axis.ticks.x  = element_blank(), axis.text.x = element_blank(),
        axis.text.y = element_text(size = 12),
        axis.title = element_text(size= 15),
        legend.text = element_text(size = 12),
        legend.title  = element_blank(),
        axis.line = element_line(size = 0.8),
        legend.position = "top") 

# ggsave("plots/HHcohort_stackedBarplot_150samples_20200522.pdf", width = 10, height = 3)
```

## Heatmap

```{r, fig.height=10, fig.width=10}
df_plot <- counts
df_plot <- df_plot[rownames(sig_matrix_new), levels(plot.data$sample)]
df_plot <- t ( scale(t(df_plot + .5), center = T, scale = T))
df_plot[df_plot >= 5 ] <- 5
df_plot[df_plot <  -2 ] <- -2
 
# pdf("plots/HHcohort_heatmap_150samples_20200522.pdf", width = 12, height = 10)
pheatmap::pheatmap(df_plot,color =  inferno(50),
                   show_colnames = FALSE,
                   scale = "none", cluster_cols = F, 
                   cluster_rows = F)
# dev.off()
```


# Technical

```{r}
sessionInfo()
```

