---
title: "Deconvolution analysis of NanoString HH-cohort data"
author: "Zhiyuan Hu"
date: "31/10/2019"
output: 
    html_document:
        toc: true
        number_sections: true
        theme: united
---


```{r setup, include=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = FALSE, fig.width = 6, fig.height = 4, warning = F)

# devtools::install_github('shenorrlab/bseq-sc')
# devtools::install_github('shenorrlab/csSAM')
library(bseqsc)

library(MASS)
library(xbioc)
library(e1071) #svm
library(preprocessCore)

library(survminer)
library(survival)
# library(powerSurvEpi)
library(Biobase)
# BiocManager::install("SingleCellExperiment")
# BiocManager::install("scater")
# library(SingleCellExperiment)
# library(scater)

library(ggbeeswarm)
library(ggrepel)
library(viridis)
library(ggplot2)
library(cowplot)
library(ggpubr)

# library(edgeR)
# library(limma)
# library(copynumber)

source("../nanostring/codings/modified_functions20190204.R")
```


# Load data

```{r}
sig_matrix_new <-  read.csv("../TableS7_selected_signature_matrix_52genes20190204.csv", as.is = T, row.names = 1)
sig_matrix_new <- as.matrix(sig_matrix_new)
```


```{r}
counts <- read.csv("../HH001-150_NormalizedData_cleaned.csv", as.is = T, row.names = 1)

rownames(counts)[7] <- "MLF1IP"

colnames(counts) <- gsub(pattern = "HH", replacement = "HH-", x = colnames(counts))
colnames(counts) <- gsub(pattern = "[.]", replacement = "", x = colnames(counts))
counts[1:5,1:5]
# counts <- read.csv("../HH_cohort/HH001-150_Rawdata_cleaned.csv", as.is = T, row.names = 1)
# counts[1:5,1:5]
```



```{r}
clic_dt <- readxl::read_xlsx("../clinical_data.xlsx")
colnames(clic_dt) <- gsub(pattern = " ", replacement = "_", x = colnames(clic_dt))
clic_dt$Overall_survival <- as.numeric(clic_dt$Overall_survival)
clic_dt$Event <- as.numeric(clic_dt$Event)
head(clic_dt)
```

```{r}
sum(rownames(counts) %in% rownames(sig_matrix_new))
```

# Deconvolution

```{r, fig.height=10}
cibersort_fit <- cibersort_my(sig_matrix = sig_matrix_new, exprs_unlogged = counts)
# cibersort_fit2 <- cibersort_my(sig_matrix = sig_matrix, exprs_unlogged = counts)

# write.csv(nanostring_fit, file = "results/cibersort_HHcohort150_output.csv")
colnames(cibersort_fit)[c(1,2,4,5)] <- c("Differentiated","KRT17","Cell_cycle","Ciliated")

clic_dt <- cbind(clic_dt, cibersort_fit[match(clic_dt$Anonymised_code, rownames(cibersort_fit)),])
```

# Filtering

```{r}
dim(clic_dt) # 150 patients
```

```{r}
hist(clic_dt$EMT, 20); abline(v = 0.95)
```


```{r filtering}
table(clic_dt$EMT < 0.95)
# FALSE  TRUE 
#     2   148 
clic_filtered <- clic_dt[clic_dt$EMT < 0.95,]
dim(clic_filtered)

table(clic_filtered$Summary_histology) 
clic_filtered <- clic_filtered[clic_filtered$Summary_histology != "Endometrial serous carcinoma",] # remove Endometrial serous carcinoma
dim(clic_filtered)

clic_filtered$Event[clic_filtered$Overall_survival == 29] <- 0 # die from compliments; 1 patient
dim(clic_filtered)

sum(is.na(clic_filtered$Summary_grade) | is.na(clic_filtered$Summary_stage)) # 19 missing grade or stage information
clic_filtered <- clic_filtered[!is.na(clic_filtered$Summary_grade) & !is.na(clic_filtered$Summary_stage),]
dim(clic_filtered)

sum(is.na(clic_filtered$Overall_survival)) # 6 dont have survival info
clic_filtered <- clic_filtered[!is.na(clic_filtered$Overall_survival),]

dim(clic_filtered)
```

# Survival analysis

## Single-variate analysis

```{r single-variate}
surv_fit <- coxph(Surv(Overall_survival, Event)~EMT, data=as.data.frame(clic_filtered))
summary(surv_fit)
```

## Multiple-variate analysis

```{r multivariate}
surv_fit <- coxph(Surv(Overall_survival, Event)~EMT+Summary_grade+Summary_stage, data=as.data.frame(clic_filtered))
summary(surv_fit)
```

```{r}
# xlsx::write.xlsx(clic_filtered,"results/clincData_deconRes_filtered20191106.xlsx")
```


```{r multivariate-res-table}
res_tb <- data.frame(Variate = c("EMT","Grade","Stage"),
                     Hazard_ratio = summary(surv_fit)$coefficients[,2],
                     p = summary(surv_fit)$coefficients[,5],
                     CI_low = summary(surv_fit)$conf.int[,3],
                     CI_high = summary(surv_fit)$conf.int[,4])
res_tb$Hazard_ratio[2] <- 1/res_tb$Hazard_ratio[2]
res_tb[2, 4:5] <- 1/res_tb[2,c(5,4)]
# write.csv(res_tb, "results/survival_Fitres_HH20191106.csv", row.names = F)
knitr::kable(res_tb)
```

## Dichotomisation

The following coding tested which cutoff of dichotomisation gives the lowest p value.

```{r}
surv_rls <- matrix(0, nrow = 1, ncol = 3)
for (i in seq(0.08, 0.57, 0.005)) {
  clic_filtered$emt_group <- NA
  clic_filtered$emt_group[clic_filtered$EMT <= i] <- "low"
  clic_filtered$emt_group[clic_filtered$EMT > i] <- "high"
  fit <- coxph(Surv(Overall_survival, Event)~emt_group, data=clic_filtered)
  surv_rls <- rbind(surv_rls, c(i,summary(fit)$coef[,2], summary(fit)$coef[,5]))
}
surv_rls <- surv_rls[-1,]
```

Ideally the best cutoff is 0.36

```{r}
plot(surv_rls[,1],surv_rls[,3], pch = 20); abline(h = 0.05)
```

## KM curves

```{r}
clic_filtered$emt_group <- NA
clic_filtered$emt_group[clic_filtered$EMT <= 0.36] <- "low"
clic_filtered$emt_group[clic_filtered$EMT > 0.36] <- "high"
fit <- coxph(Surv(Overall_survival, Event)~emt_group+Summary_stage, data=clic_filtered)
summary(fit)
```

```{r}
table(clic_filtered$emt_group)
```

28 out of 122 (i.e. 23% patients are classified as EMT-high)

```{r kmcurve-earlystage, fig.height=6, fig.width=4.5}
fit1 <- survfit(Surv(Overall_survival, Event)~emt_group, data = clic_filtered[clic_filtered$Summary_stage == "early",])
# ggsurvplot(fit1, data = clic_filtered, pval = T, risk.table = F)
# ggsave("plots/HHcohort_earlyStage_EMTgroup_KMcurve20200122.pdf", width = 6,height = 5)

ggsurvplot(fit1, data = clic_filtered, pval = T, risk.table = T)
# ggsave("plots/HHcohort_earlyStage_EMTgroup_riskTable20200122.pdf", width = 6,height = 2)
```

```{r km-curve-latestage, fig.height=6, fig.width=4.5}
fit1 <- survfit(Surv(Overall_survival, Event)~emt_group, data = clic_filtered[clic_filtered$Summary_stage == "late",])
# ggsurvplot(fit1, data = clic_filtered, pval = T, risk.table = F)
# ggsave("plots/HHcohort_lateStage_EMTgroup_KMcurve20200122.pdf", width = 6,height = 5)

ggsurvplot(fit1, data = clic_filtered, pval = T, risk.table = T)
# ggsave("plots/HHcohort_lateStage_EMTgroup_riskTable20200122.pdf", width = 6,height = 2)
```


# Compare Imperial, TCGA and AOCS

```{r}
path <- "~/OneDrive - OnTheHub - The University of Oxford/Project scRNAseq analysis/A_SmartSeq2/man_analysis2_20180629/R_deconvolution/rds"
tcga_eset <- readRDS(paste(path, "/20181029TCGA_eset.rds", sep = ""))
aocs_eset <- readRDS(paste(path, "/20181029_tothill_eset.rds", sep = ""))
aocs_dec <- readRDS("~/OneDrive - OnTheHub - The University of Oxford/Project scRNAseq analysis/A_SmartSeq2/revision_analysis_20190827/R_deconvolution/results/20190502Deconvolution_Tothill.rds")
cod_esets <- readRDS(paste(path, "/20190213CuratedOvarianData_esets.rds", sep = ""))
# write.table(na.omit(expm1(cod_esets[[7]]@assayData$exprs)), row.names = T, file = "../CuratedOvarianData/GSE51088_expm1_exprs.txt", sep = "\t", quote = F)
```

```{r clean-public-datasets}
aocs_eset$OS.time <- aocs_eset$time.to.death
aocs_eset$OS <- NA
aocs_eset$OS[aocs_eset$patient.status == "D"] <- 1
aocs_eset$OS[aocs_eset$patient.status %in% c("D*","PF","R")] <- 0

cod_info <- rbind(cod_esets[[1]]@phenoData@data,
                  cod_esets[[2]]@phenoData@data,
                  cod_esets[[3]]@phenoData@data,
                  cod_esets[[4]]@phenoData@data,
                  cod_esets[[5]]@phenoData@data,
                  cod_esets[[6]]@phenoData@data,
                  cod_esets[[7]]@phenoData@data)

sapply(cod_esets, ncol)

database <- list("E.MTAB.386", "GSE13876", "GSE26193",
              "GSE26712", "GSE32062.GPL6480",
              "GSE49997", "GSE51088")

cod_info$dataset <- unlist(mapply(database, sapply(cod_esets, ncol), FUN = function(x,y) return(rep(x = x, y))))
cod_info <- cod_info[cod_info$sample_type %in% c("tumor","metastatic"),]
cod_info <- cod_info[!is.na(cod_info$vital_status),]
cod_info$os_binary[cod_info$vital_status == "living"] <- 0
cod_info$os_binary[cod_info$vital_status == "deceased"] <- 1


aocs_eset$summary_stage <- NA
aocs_eset$summary_stage[aocs_eset$Stage %in% c("I","IA","IB","IC","IIA","IIB","IIC")] <-"early"
aocs_eset$summary_stage[aocs_eset$Stage %in% c("III","IIIA","IIIB","IIIC","IV")] <- "late"
aocs_eset$EMT <- aocs_dec[,"EMT"]

tcga_eset$stage <- NA
tcga_eset$stage[tcga_eset$clinical_stage %in% c("Stage IA", "Stage IB", "Stage IC", "Stage IIA", "Stage IIB", "Stage IIC")] <- "early"
tcga_eset$stage[tcga_eset$clinical_stage %in% c("Stage IIIA", "Stage IIIB", "Stage IIIC", "Stage IV")] <- "late"
```


```{r, fig.width=6, fig.height=5}
df <- data.frame(OS.time = c(clic_filtered$Overall_survival, tcga_eset$OS.time, aocs_eset$OS.time*30, cod_info$days_to_death),
                 OS = c(clic_filtered$Event, tcga_eset$OS, aocs_eset$OS, cod_info$os_binary),
                 # emt_new = c(clic_dt$emt_new,eset$emt_new, aocs_eset$emt_new),
                 dataset = c(rep("HH",nrow(clic_filtered)), rep("TCGA", ncol(tcga_eset)), rep("AOCS", ncol(aocs_eset)), cod_info$dataset),
                 stage = c(clic_filtered$Summary_stage, tcga_eset$stag, aocs_eset$summary_stage, cod_info$summarystage))
fit1 <- survfit(Surv(OS.time, OS)~dataset, df[df$stage == "late",])
ggsurvplot(fit1, df[df$stage == "late",], pval = T, risk.table = F )

# ggsave("plots/compare_OS_acrossDataset20200122.pdf", width = 12, height = 9)
```

```{r, fig.width=6, fig.height=5}
df <- df[1:715,]
fit1 <- survfit(Surv(OS.time, OS)~dataset, df[df$stage == "late",])
ggsurvplot(fit1, df[df$stage == "late",], pval = T, risk.table = F )
# ggsave("plots/compare_OS_acrossDataset_tcga_aocs20200122.pdf", width = 8, height = 6)
```

# Low-grade = ciliated

```{r Low-grade-ciliated, fig.height=4, fig.width=3}
ggplot(clic_filtered, aes(x = Summary_grade, y = Ciliated)) + 
    geom_violin(scale = "width") + geom_boxplot(width = 0.1) + geom_jitter(alpha = 0.3) + 
    ylab("Ciliated scores") + xlab("Grade") + theme_classic2()
# ggsave("plots/compare_lowHighGrade20200122.pdf", width = 3, height = 3)
```

# Visualisation

## Stacked barplot

```{r Stacked barplot-preparedata}
svr_fit_oxo <- clic_filtered[,25:29]
svr_fit_new_diff <- svr_fit_oxo[svr_fit_oxo[,1] > 0.5,]
svr_fit_new_diff <- svr_fit_new_diff[order(svr_fit_new_diff[,1], decreasing = F),]
svr_fit_new_krt <- svr_fit_oxo[svr_fit_oxo[,2] > 0.5,]
svr_fit_new_krt <- svr_fit_new_krt[order(svr_fit_new_krt[,2], decreasing = F),]
svr_fit_new_emt <- svr_fit_oxo[svr_fit_oxo[,3] > 0.5,]
svr_fit_new_emt <- svr_fit_new_emt[order(svr_fit_new_emt[,3], decreasing = F),]
svr_fit_new_cc <- svr_fit_oxo[svr_fit_oxo[,4] > 0.5,]
svr_fit_new_cc <- svr_fit_new_cc[order(svr_fit_new_cc[,4], decreasing = F),]

svr_fit_new_others <- svr_fit_oxo[!rownames(svr_fit_oxo) %in% c(rownames(svr_fit_new_diff),rownames(svr_fit_new_krt),
                                                               rownames(svr_fit_new_emt),rownames(svr_fit_new_cc)),]

plot.data <- data.frame(sample = rep(rownames(svr_fit_oxo), 5),
                        state = rep(c("Differentiated","KRT17","EMT","Cell cycle","Ciliated"), 
                                    each = nrow(svr_fit_oxo)),
                        score = c(svr_fit_oxo[,1],
                                  svr_fit_oxo[,2],
                                  svr_fit_oxo[,3],
                                  svr_fit_oxo[,4],
                                  svr_fit_oxo[,5]))
plot.data$sample <- factor(plot.data$sample, levels = c(rownames(svr_fit_new_diff),rownames(svr_fit_new_krt),
                                                               rownames(svr_fit_new_emt),rownames(svr_fit_new_cc),
                                                        rownames(svr_fit_new_others)))
plot.data$state <- factor(plot.data$state, levels = c("Differentiated","KRT17","EMT","Cell cycle","Ciliated"))
# differentiated #A16BA3
# F6A000 KRT7
# FFDE00 EMT
# CC #4DD9FF
#F3A1B6 Ciliated
```

```{r, fig.width=10, fig.height=3}
ggplot(plot.data) + geom_bar(aes(y = score, x = sample, fill = state),
                           stat="identity") + theme_classic() + 
  scale_fill_manual(values = c("#A16BA3","#F6A000","#FFDE00","#4DD9FF","#F3A1B6"), 
                    breaks = c("Differentiated","KRT17","EMT","Cell cycle","Ciliated"), 
                    labels = c("Differentiated","KRT17","EMT","Cell cycle","Ciliated")) +
  xlab("Bulk tumour samples \n (each column is one tumour case)") + ylab("Proportions of cell states") +
  theme(axis.ticks.x  = element_blank(), axis.text.x = element_blank(),
        axis.text.y = element_text(size = 12),
        axis.title = element_text(size= 15),
        legend.text = element_text(size = 12),
        legend.title  = element_blank(),
        axis.line = element_line(size = 0.8)) 

# ggsave("plots/HHcohort_stackedBarplot_20200122.pdf", width = 10, height = 3)
```

## PCA plot

```{r, fig.width=10, fig.height=7}
m_pc <- prcomp(t(log1p(counts[,match(rownames(clic_filtered), colnames(counts))])))

clic_filtered$PC1 <- m_pc$x[,1]
clic_filtered$PC2 <- m_pc$x[,2]


p1 <- ggplot(clic_filtered, aes(x = PC1, y = PC2, fill = EMT)) + geom_point(size = 2, pch = 21, col = "grey") + theme_classic() + 
    scale_fill_viridis(option="inferno") + theme(axis.title = element_blank() , legend.text = element_text(size = 7))
p2 <- ggplot(clic_filtered, aes(x = PC1, y = PC2, fill = Cell_cycle)) + geom_point(size = 2, pch = 21, col = "grey") + theme_classic() + 
    scale_fill_viridis(option="inferno") + theme(axis.title = element_blank() , legend.text = element_text(size = 7))
p3 <- ggplot(clic_filtered, aes(x = PC1, y = PC2, fill = Differentiated)) + geom_point(size = 2, pch = 21, col = "grey") + theme_classic() + 
    scale_fill_viridis(option="inferno") + theme(axis.title = element_blank() , legend.text = element_text(size = 7))
p4 <- ggplot(clic_filtered, aes(x = PC1, y = PC2, fill = KRT17)) + geom_point(size = 2, pch = 21, col = "grey") + theme_classic() + 
    scale_fill_viridis(option="inferno") + theme(axis.title = element_blank() , legend.text = element_text(size = 7))
p5 <- ggplot(clic_filtered, aes(x = PC1, y = PC2, fill = Ciliated)) + geom_point(size = 2, pch = 21, col = "grey") + theme_classic() + 
    scale_fill_viridis(option="inferno") + theme(axis.title = element_blank() , legend.text = element_text(size = 7))
cowplot::plot_grid(p1,p2,p3,p4,p5, ncol = 3)
# ggsave("plots/HHcohort_PCAplots_decovScores20200122.pdf", width = 10, height = 7)
```

# TCGA analysis

```{r ca}
tcga_fit <- read.csv("data/20191125TCGA_deconvolutionResults.csv", row.names = 1)
```

## LM22

```{r}
immune <- read.delim("data/CIBERSORT.Output_Job18.txt", row.names = 1)
immune <- immune[match(rownames(tcga_fit), rownames(immune)),]
```

```{r}
LM22 <- read.delim("data/LM22.txt", as.is = T, row.names = 1)
LM22_macrophage_markers <- names(which(apply(LM22, 1, which.max) == 16))
LM22_macrophage_markers
```

### Correlation analysis

```{r}
cor.rls <- t(cor(tcga_fit[,1:5], immune[,1:22]))
# pdf("plots/TCGA_cortest_LM22_20200122.pdf")
pheatmap::pheatmap(cor.rls, display_numbers = round(cor.rls,1), cellheight = 12, cellwidth = 30)
# dev.off()
```

```{r}
LM22_macrophage_markers <- unique(c(LM22_macrophage_markers, "CD163","TGM2"))

cor_res <- c()
for(itor_gene in LM22_macrophage_markers){
    if(itor_gene %in% rownames(tcga_eset)){
        
        tmp <- cor(tcga_fit[,"EMT"], log1p(tcga_eset@assayData$exprs[itor_gene,]))
        cor_res <- rbind(cor_res, c(itor_gene, tmp))
    }
}
cor_res <- data.frame(gene = cor_res[,1],
                      cor = as.numeric(cor_res[,2]))
cor_res <- cor_res[order(cor_res$cor, decreasing = T),]
cor_res$sign <- cor_res$cor > 0
cor_res$gene <- factor(cor_res$gene, levels = cor_res$gene)
```


The expression levels of most of the macrophage M2 markers have a positive correlation with the EMT score.

```{r correlation_plot, fig.width=8, fig.height=3}
ggplot(cor_res, aes(x = gene, y = cor, fill = sign)) + geom_bar(stat = "identity") + theme_classic2() + theme(legend.position = "none", axis.text.x = element_text(angle = 30, vjust = 0.5, hjust = 0.5)) + xlab("") + ylab("Correlation with EMT scores")

# ggsave("plots/TCGAcorrelation_barplot_M2markers20200122.pdf", width = 8, height = 5)
```



```{r}
cor(tcga_eset@assayData$exprs["SPARC",], tcga_eset@assayData$exprs["CD163",])
```

### Linear regression

linear relationship exists between macrophage M2 marker CD163 and EMT marker SPARC

```{r}
lr_res <- lm(log10(tcga_eset@assayData$exprs["CD163",]+1) ~log10(tcga_eset@assayData$exprs["SPARC",]+1))
summary(lr_res)
```

```{r}
df_plot <- data.frame(SPARC = log10(tcga_eset@assayData$exprs["SPARC",] + 1),
                      CD163 = log10(tcga_eset@assayData$exprs["CD163",] + 1))

ggplot(df_plot, aes(x = SPARC, y = CD163))  + geom_abline(slope = lr_res$coefficients[2], intercept = lr_res$coefficients[1], col = "red")+ geom_point(alpha = .6) + theme_linedraw() + xlim(3,6) + ylim(1,4)
# ggsave("plots/TCGA_linearRes_scatter_CD163vsSPARC20200122.pdf", width = 5, height = 5)
```


```{r}
# res <- c()
# for(i in 1:nrow(tcga_eset@assayData$exprs)){
#     tmp <- lm(log10(tcga_eset@assayData$exprs["CD163",] +1)~log10(tcga_eset@assayData$exprs[i,] +1))
#     res <- cbind(res, tmp$coefficients[2])
# }
# res
```

### t test

```{r}
df <- cbind(tcga_fit[,1:5], immune)
df$EMT.group <- "EMT-middle"
df$EMT.group[df$EMT < quantile(df$EMT, 1/3)] <- "EMT-low"
df$EMT.group[df$EMT > quantile(df$EMT, 2/3)] <- "EMT-high"
df$EMT.group <- factor(df$EMT.group, levels = c("EMT-low", "EMT-middle", "EMT-high"))

df$KRT17.group <- "KRT17-middle"
df$KRT17.group[df$KRT17 <= quantile(df$KRT17, 1/3)] <- "KRT17-low"
df$KRT17.group[df$KRT17 > quantile(df$KRT17, 2/3)] <- "KRT17-high"
df$KRT17.group <- factor(df$KRT17.group, levels = c("KRT17-low", "KRT17-middle", "KRT17-high"))
```


```{r}
ggplot(df, aes(x = EMT.group, y =  Macrophages.M2)) + geom_violin(scale = "area") + 
    geom_boxplot(width = 0.2) + geom_beeswarm(alpha = 0.4) + theme_classic2()
# ggsave("plots/TCGA_EMTgroup_versus_macrophageM2.pdf", width = 4, height = 5)
```

```{r}
t.test(df$Macrophages.M2[df$EMT.group == "EMT-high"], df$Macrophages.M2[df$EMT.group == "EMT-low"], alternative = "greater")
```

## LM6

```{r}
immune2 <- read.delim("data/CIBERSORT.Output_Job19.txt", row.names = 1)
immune2 <- immune2[match(rownames(tcga_fit), rownames(immune2)),]
```

### Correlation test

```{r}
cor.rls2 <- t(cor(tcga_fit[,1:5], immune2[,1:6]))
# pdf("plots/TCGA_cortest_LM6_20200122.pdf")
pheatmap::pheatmap(cor.rls2, display_numbers = round(cor.rls2,1), cellheight = 12, cellwidth = 30)
# dev.off()
```

### t-test

```{r}
df <- cbind(tcga_fit[,1:5], immune2)
df$EMT.group <- "EMT-middle"
df$EMT.group[df$EMT < quantile(df$EMT, 1/3)] <- "EMT-low"
df$EMT.group[df$EMT > quantile(df$EMT, 2/3)] <- "EMT-high"
df$EMT.group <- factor(df$EMT.group, levels = c("EMT-low", "EMT-middle", "EMT-high"))

df$KRT17.group <- "KRT17-middle"
df$KRT17.group[df$KRT17 <= quantile(df$KRT17, 1/3)] <- "KRT17-low"
df$KRT17.group[df$KRT17 > quantile(df$KRT17, 2/3)] <- "KRT17-high"
df$KRT17.group <- factor(df$KRT17.group, levels = c("KRT17-low", "KRT17-middle", "KRT17-high"))
```

```{r}
ggplot(df, aes(x = EMT.group, y =  Monocytes)) + geom_violin(scale = "area") + 
    geom_boxplot(width = 0.2) + geom_beeswarm(alpha = 0.4) + theme_classic2()
 # ggsave("plots/TCGA_EMTgroup_versus_monocytes.pdf", width = 4, height = 5)
```

```{r}
t.test(df$Monocytes[df$EMT.group == "EMT-high"], df$Monocytes[df$EMT.group == "EMT-low"], alternative = "greater")
```

## Survival analysis

```{r}
tcga_eset$EMT <- tcga_fit$EMT[match(colnames(tcga_eset), rownames(tcga_fit))]
tcga_eset$MacrophageM2 <- immune$Macrophages.M2[match(colnames(tcga_eset), rownames(immune))]
tcga_eset$Monocyte <- immune2$Monocytes[match(colnames(tcga_eset), rownames(immune2))]
df <- data.frame(OS = tcga_eset$OS,
                 OS.time = tcga_eset$OS.time,
                 EMT = tcga_eset$EMT,
                 stage = as.factor(tcga_eset$stage),
                 MacrophageM2 = tcga_eset$MacrophageM2,
                 Monocyte = tcga_eset$Monocyte,
                 CD163 = tcga_eset@assayData$exprs["CD163",],
                 CD209 = tcga_eset@assayData$exprs["CD209",],
                 SPARC = tcga_eset@assayData$exprs["SPARC",])
df$CD163b <- NA
df$CD163b[df$CD163 > quantile(df$CD163, 0.8)] <- "CD163-high"
df$CD163b[df$CD163 <= quantile(df$CD163, 0.8)] <- "CD163-low"

df$EMTb <- NA
df$EMTb[df$EMT > quantile(df$EMT, 0.5)] <- "EMT-high"
df$EMTb[df$EMT <= quantile(df$EMT, 0.5)] <- "EMT-low"


rownames(df) <- colnames(tcga_eset)
surv_fit <- coxph(Surv(tcga_eset$OS.time,tcga_eset$OS)~ df$EMT + df$CD163 + df$stage)
summary(surv_fit)
```

```{r}
surv_fit <- coxph(Surv(tcga_eset$OS.time,tcga_eset$OS)~ df$EMT + df$MacrophageM2 + df$stage)
summary(surv_fit)
```

Macrophage propotions slightly correlate with prognosis in TCGA.

```{r}
surv_fit <- coxph(Surv(tcga_eset$OS.time,tcga_eset$OS)~ df$MacrophageM2 + df$stage)
summary(surv_fit)
```



```{r}
chisq.test(table(df$CD163b, df$EMTb))
```

```{r}
cor.test(df$CD163, df$EMT)
```

## CD163 & EMT

A addictive effect observed in TCGA data.

```{r}
fit1 <- survfit(Surv(df$OS.time,df$OS)~df$CD163b+df$EMTb)
ggsurvplot(fit1, df, pval = T, risk.table = T)
```


```{r}
cor.test(df$CD163, df$Monocyte)
```

```{r}
cor.test(df$CD163, df$MacrophageM2)
```

# AOCS

## Cibersort result

```{r}
aocs_lm22 <- read.csv("results/CIBERSORT.Output_Job20_AOCS_LM22.csv")
aocs_eset$macrophageM2 <- aocs_lm22$Macrophages.M2

aocs_lm6 <- read.csv("results/CIBERSORT.Output_Job21_AOCS_LM6.csv")
aocs_eset$monocyte <- aocs_lm6$Monocytes
```


## Correlation

```{r}
cor.rls <- t(cor(aocs_dec[,1:5], aocs_lm22[,2:23]))
# pdf("plots/AOCS_cortest_LM22_20200131.pdf")
pheatmap::pheatmap(cor.rls, display_numbers = round(cor.rls,1), cellheight = 12, cellwidth = 30)
# dev.off()
```


```{r}
cor.rls <- t(cor(aocs_dec[,1:5], aocs_lm6[,2:7]))
# pdf("plots/TCGA_cortest_LM6_20200131.pdf")
pheatmap::pheatmap(cor.rls, display_numbers = round(cor.rls,1), cellheight = 12, cellwidth = 30)
# dev.off()
```

## t-test

```{r}
df <- cbind(aocs_dec[,1:5],aocs_lm22[,1:23])
df$EMT.group <- "EMT-middle"
df$EMT.group[df$EMT <= quantile(df$EMT, 1/3)] <- "EMT-low"
df$EMT.group[df$EMT > quantile(df$EMT, 2/3)] <- "EMT-high"
df$EMT.group <- factor(df$EMT.group, levels = c("EMT-low", "EMT-middle", "EMT-high"))

ggplot(df, aes(x = EMT.group, y =  Macrophages.M2)) + geom_violin(scale = "width") + 
    geom_boxplot(width = 0.1) + geom_jitter(alpha = 0.4) + theme_classic2()
# ggsave("plots/AOCS_EMTgroup_versus_MacrophageM2.pdf", width = 4, height = 5)
```

```{r}
t.test(df$Macrophages.M2[df$EMT.group == "EMT-high"], df$Macrophages.M2[df$EMT.group == "EMT-low"], alternative = "greater")
```

```{r}
df <- cbind(aocs_dec[,1:5],aocs_lm6[,1:6])
df$EMT.group <- "EMT-middle"
df$EMT.group[df$EMT <= quantile(df$EMT, 1/3)] <- "EMT-low"
df$EMT.group[df$EMT > quantile(df$EMT, 2/3)] <- "EMT-high"
df$EMT.group <- factor(df$EMT.group, levels = c("EMT-low", "EMT-middle", "EMT-high"))

ggplot(df, aes(x = EMT.group, y =  Monocytes)) + geom_violin(scale = "width") + 
    geom_boxplot(width = 0.1) + geom_jitter(alpha = 0.4) + theme_classic2()
# ggsave("plots/AOCS_EMTgroup_versus_Monocytes.pdf", width = 4, height = 5)
```

```{r}
t.test(df$Monocytes[df$EMT.group == "EMT-high"], df$Monocytes[df$EMT.group == "EMT-low"], alternative = "greater")
```

Difference in the correlation test: technical issue?

## Survival

```{r}
surv_fit <- coxph(Surv(aocs_eset$OS.time,aocs_eset$OS)~ aocs_eset$EMT + aocs_eset$summary_stage)
summary(surv_fit)
```

CD163 is no longer associated with survival in AOCS data

Where the TAMs are also matters? See https://ovarianresearch.biomedcentral.com/articles/10.1186/1757-2215-7-19

Macrophage M2 proportions slightly correlates with survival in AOCS too.

```{r}
surv_fit <- coxph(Surv(aocs_eset$OS.time,aocs_eset$OS)~ aocs_eset$macrophageM2 + aocs_eset$summary_stage)
summary(surv_fit)
```

# CuratedOvarianCancer

```{r}

```


# Technical

```{r}
sessionInfo()


# R version 3.6.1 (2019-07-05)
# Platform: x86_64-apple-darwin15.6.0 (64-bit)
# Running under: macOS Mojave 10.14.3
# 
# Matrix products: default
# BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib
# LAPACK: /Library/Frameworks/R.framework/Versions/3.6/Resources/lib/libRlapack.dylib
# 
# locale:
# [1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8
# 
# attached base packages:
# [1] stats4    parallel  stats     graphics  grDevices utils     datasets  methods   base     
# 
# other attached packages:
#  [1] RColorBrewer_1.1-2    ggpubr_0.2.1          magrittr_1.5          cowplot_0.9.4         viridis_0.5.1        
#  [6] viridisLite_0.3.0     ggrepel_0.8.1         ggbeeswarm_0.6.0      ggplot2_3.2.1         survival_2.44-1.1    
# [11] preprocessCore_1.46.0 e1071_1.7-2           xbioc_0.1.17          AnnotationDbi_1.46.1  IRanges_2.18.3       
# [16] S4Vectors_0.22.1      MASS_7.3-51.4         bseqsc_1.0            Biobase_2.44.0        BiocGenerics_0.30.0  
# [21] csSAM_1.4             Rcpp_1.0.2           
# 
# loaded via a namespace (and not attached):
#  [1] tsne_0.1-3         fs_1.3.1           usethis_1.5.0      devtools_2.0.2     bit64_0.9-7        doParallel_1.0.15 
#  [7] rprojroot_1.3-2    tools_3.6.1        backports_1.1.5    utf8_1.1.4         R6_2.4.0           vipor_0.4.5       
# [13] DBI_1.0.0          lazyeval_0.2.2     colorspace_1.4-1   withr_2.1.2        tidyselect_0.2.5   gridExtra_2.3     
# [19] prettyunits_1.0.2  processx_3.3.1     bit_1.1-14         curl_3.3           compiler_3.6.1     cli_1.1.0         
# [25] desc_1.2.0         pkgmaker_0.28      labeling_0.3       scales_1.0.0       callr_3.2.0        NMF_0.22          
# [31] stringr_1.4.0      digest_0.6.22      pkgconfig_2.0.3    bibtex_0.4.2       sessioninfo_1.1.1  limma_3.40.6      
# [37] readxl_1.3.1       rlang_0.4.1        rstudioapi_0.10    RSQLite_2.1.2      dendextend_1.12.0  dplyr_0.8.3       
# [43] Formula_1.2-3      Matrix_1.2-17      fansi_0.4.0        munsell_0.5.0      stringi_1.4.3      yaml_2.2.0        
# [49] edgeR_3.26.8       pkgbuild_1.0.3     plyr_1.8.4         grid_3.6.1         blob_1.2.0         crayon_1.3.4      
# [55] lattice_0.20-38    splines_3.6.1      locfit_1.5-9.1     zeallot_0.1.0      knitr_1.23         ps_1.3.0          
# [61] pillar_1.4.2       ggsignif_0.5.0     rngtools_1.4       reshape2_1.4.3     codetools_0.2-16   pkgload_1.0.2     
# [67] glue_1.3.1         remotes_2.1.0      BiocManager_1.30.9 vctrs_0.2.0        foreach_1.4.7      cellranger_1.1.0  
# [73] gtable_0.3.0       purrr_0.3.3        assertthat_0.2.1   xfun_0.8           gridBase_0.4-7     xtable_1.8-4      
# [79] class_7.3-15       pheatmap_1.0.12    tibble_2.1.3       iterators_1.0.12   beeswarm_0.2.3     registry_0.5-1    
# [85] memoise_1.1.0      cluster_2.1.0  
```
